\name{}: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-04-18 paper/main.tex
.

Both bottom-up and top-down approaches to drug--target interaction
prediction have some advantages and limitations. Generally, bottom-up
methods face the challenge to predict whether a structure binds to a
protein given their molecular properties; whether two entities
interact depends not only on the molecular structure of the entities
(where binding sites and molecular forces need to be determined for
accurate prediction) but also properties such as where a protein is
expressed or in which celltypes and anatomical structures. Top-down
methods use information for DTI prediction such as side-effect
similarity \cite{Campillos2008} that is largely complementary to
knowledge gained from molecular properties. While methods tha3t rely on
molecular information are directly predicting whether two molecular
could interact, top-down methods rely on more indirect means and 
infer a DTI from observable effects resulting from the interaction.

Both approaches may be used in conjunction with network inference
\todo[inline]{CITE}\cite{}. Biological networks used for DTI prediction include
protein--protein interaction networks \todo[inline]{CITE}\cite{} and networks including
several other types of biological relations, including similarity
between represented entities \cite{Gottlieb2011}. Network-based DTI
prediction methods use the guilt-by-association principle
\cite{Oliver2000}
and assume that a protein is a likely target for a drug if many of the
protein's neighbors in the interaction network are targets of the drug
\cite{Gillis2012}. Network-based methods have been applied successfully
to DTI prediction. However, if DTIs are taken as direct physical
interactions between a drug and protein, it remains an unresolved
question whether the network-based guilt-by-association hypothesis is
true, or whether an interaction of a drug and protein dys-regulates
several of the protein's interaction partners, and therefore resulting
in downstream effects that are not direct interactions but only
downstream consequences of an interaction.

>>>>

Both bottom-up and top-down approaches to drug--target interaction
prediction have some advantages and limitations. Generally, bottom-up
methods face the challenge to predict whether a structure binds to a
protein given their molecular properties; whether two entities
interact depends not only on the molecular structure of the entities
(where binding sites and molecular forces need to be determined for
accurate prediction) but also properties such as where a protein is
expressed or in which celltypes and anatomical structures. Top-down
methods use information for DTI prediction such as side-effect
similarity \cite{Campillos2008} that is largely complementary to
knowledge gained from molecular properties. While methods that rely on
molecular information are directly predicting whether two molecular
could interact, top-down methods rely on more indirect means and 
infer a DTI from observable effects resulting from the interaction.

Both approaches may be used in conjunction with network inference
\cite{Chen2015}. Biological networks used for DTI prediction include
protein--protein interaction networks \cite{Feng2017, Lee2018} and
networks including several other types of biological relations,
including similarity between represented entities
\cite{Gottlieb2011}. Network-based DTI prediction methods use the
guilt-by-association principle \cite{Oliver2000} and assume that a
protein is a likely target for a drug if many of the protein's
neighbors in the interaction network are targets of the drug
\cite{Gillis2012}. Network-based methods have been applied
successfully to DTI prediction. However, if DTIs are taken as direct
physical interactions between a drug and protein, it remains an
unresolved question whether the network-based guilt-by-association
hypothesis is true, or whether an interaction of a drug and protein
dys-regulates several of the protein's interaction partners, and
therefore resulting in effects that are not direct interactions but
only downstream consequences of an interaction.
